Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-25 src/8.Related.tex
Adding other example and back-end example.
Updated Conclusion

\section{Related Work}\label{chap:related}
Like Kokkos, RAJA~\cite{RAJA} is a C++ framework for porting massive DOE codes to a variety of architectures. In both frameworks, a developer first expresses the operations their kernel performs using simple C++ functional programming mechanisms. The difference is that Kokkos is declarative, while RAJA is descriptive. In Kokkos, the framework figures out how to map that kernel to different architectures. In RAJA, functionality is provided to make it easy for a developer to try different strategies to map those kernels to architectures, but the developer still decides how that mapping happens. Which model makes more sense in a given circumstance depends on how much exposure to the underlying programming model a developer wants. RAJA is a viable option for class education in which the educator wants to expose students to the semantics of different programming models without exposing them to the detailed syntax of those models.

@@

C++ increasingly supports parallel constructs\cite{CPP}. We have highlighted contributions of Kokkos to C++ in~\ref{chap:c++}. It is becoming increasingly possible to teach parallelism in C++ and relying on modern C++ compilers. However, at time of publication, C++ does not support accelerators, or many of the complexities of parallel computing related to accelerator programming. In courses that do not target accelerators, this can be a relevant option.

>>>>

\section{Related Work}\label{chap:related}
RAJA~\cite{RAJA} is a C++ abstraction library with similarities to Kokkos. The important different however is that Kokkos is descriptive, while RAJA is prescriptive. In Kokkos, the framework determines how an application is mapped to the underlying hardware. RAJA provides functionality that exposes hardware details and relies on the developer to try different strategies to map kernels to architectures. RAJA is a viable option for class education in which the educator would like to expose students to the semantics of different programming models without exposing them to the detailed syntax of those models.

@@

C++ increasingly supports parallel constructs. We have highlighted contributions of Kokkos to C++ in~\ref{chap:c++}. It is becoming increasingly possible to teach parallelism in C++ and relying on modern C++ compilers. However, at time of publication, C++ does not support accelerators, or many of the complexities of parallel computing related to accelerator programming. In courses that do not target accelerators, this can be a relevant option.

### 2020-01-25 src/9.Conclusion.tex
Minor bug.
Updated Conclusion

\section{Conclusion}\label{chap:conclusion}
%...
%Consequently, we emphasize the importance and welcome new development and educational use of parallel programming models that align to this trend. We believe that abstracting from vendor or architecture-specific programming models not only simplifies parallel programming in science and industry but also would support educators in teaching parallel programming.
%We would also like to express gratitude to educators for taking on this important role. Further materials, including a detailed description of the programming- and machine model can be found by following the reference on the Kokkos programming mode. Finnaly, we would like to mention that interested developers and student can reach the development team on Slack by following the invititation link as mentioned in the reference section.

>>>>

\section{Conclusion}\label{chap:conclusion}
In this work we gave an introduction to the Kokkos parallel programming model and presented guiding thoughts behind its design. We discussed a machine representation of a parallel hardware architecture that is characterized by a variety of processing units, memories and their hierarchical organization. This machine model allowed us to define a constrained semantic represented though different types and patterns. Further we listed supported paradigms and considerations that influenced the design of the semantic capture.

An example application gave insight into parallel programming with patterns and into the back-end implementation. We used them to show the generic programming interface and type specialization of the implementing classes to support a particular vendor programming model.

To the educational community, we presented Kokkos as a representative programming model of a trend towards support of parallel programming in base languages and libraries. We believe in the importance of embracing this trend in computer science education as such parallel programming models can contribute to the generic and relevant education of parallel programming and prepare student for their programming careers.

Finally we would like to express gratitude to educators for taking on this important role. We are open to provide further material or support and propose the interested reader to reach out to the development team on Slack\cite{KOKKOS_SLACK}.
